---
title: "hw2-Rohit-Thakur"
author: "Rohit Thakur"
date: "1/24/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown
Problem 1 ->

Find a dataset that is personally interesting to you. 
It may be a publicly-available dataset, or a dataset for
which you have permission to use and share results.
There are many places online to find publicly-available
dataset, and simply searching Google for your preferred topic plus “public dataset” may provide many hits.
Here some additional resources to get you started:
• US Government datasets (https://catalog.data.gov/dataset)
• Center for Disease Control (CDC) data (https://data.cdc.gov)
• Bureau of Labor Statistics (https://www.bls.gov/data/)
• NASA datasets (https://nssdc.gsfc.nasa.gov)
• World Bank Open Data (https://data.worldbank.org)
• Kaggle Datasets (https://www.kaggle.com/datasets)
This does not have to be the same dataset you will use for your group project.
Import the dataset into R, put it into a tidy format, and print the first ten observations of the dataset

Answer:

```{r}
library(dplyr)
library(tidyr)
df<-read.csv("C:/Users/rthak/OneDrive/Desktop/crime.csv")
head(df,n=10)
df1<-select(df,-c("INCIDENT_NUMBER","OFFENSE_CODE","Location","OFFENSE_DESCRIPTION"))
df1<-separate(df1,col="OCCURRED_ON_DATE",into=c("Date","Time"),sep=" ") 
df1<-filter(df1,df1$DISTRICT!="")

```

Problem 2 -> 

Perform exploratory data analysis on the dataset, using the techniques learned in class. Calculate
summary statistics that are of interest to you and create plots using ggplot2 that show your findings.

Answer:

Observations made from below graph:
UCR_PART three crimes are largest because they are non_violent 
and most occuring incidents(for ex. medical emergency).
We will further investigate insights for part one crimes as they are most violent ones.
```{r}
library(ggplot2)
df1<-filter(df1,UCR_PART!=" ")
levels(df1$UCR_PART)
df1$UCR_PART<-droplevels(df1$UCR_PART,exclude="")
df1<-na.omit(df1)
ggplot(df1)+geom_bar(aes(x=UCR_PART),stat='count')
```

Focusing on Part one crimes
```{r}
part1<-subset(df1,UCR_PART=="Part One")
```


```{r}
bypart<-part1%>%group_by(OFFENSE_CODE_GROUP)%>%summarise(occurence=n())
bypart
```

UCR PART ONE CRIMES

```{r}
ggplot(bypart)+geom_bar(aes(reorder(x=OFFENSE_CODE_GROUP,occurence),y=occurence,fill=OFFENSE_CODE_GROUP),stat="identity")+theme(axis.text=element_text(size=5),axis.title=element_text(size=8,face="bold"),legend.position = "top")+scale_color_brewer()+labs(x="OFFENSE")
```
UCR PART_1 crime by district: D4 has highest number of crimes. D4 can be categorized as unsafe district.
```{r}
ggplot(part1)+geom_bar(aes(x=DISTRICT,fill="red"))+geom_text(aes(x=DISTRICT,label=stat(count)),stat='count',nudge_y=1000)
```


UCR_PART by day of week
interesting observations: For every type of UCR_PART Fridays are having more crimes than other days.

```{r}
ggplot(df1)+geom_histogram(aes(x=UCR_PART,fill=DAY_OF_WEEK),stat='count',position="dodge")
```
From bar graph below, larceny has greatest share in d4 crimes.
```{r}
d4<-subset(part1,part1$DISTRICT==c("D4","B2","A1"))
ggplot(d4)+geom_histogram(aes(x=OFFENSE_CODE_GROUP),stat='count')+theme(axis.text.x = element_text(size=8,angle = 90))+facet_grid(~DISTRICT)
```

Problems 3–4 use data from the Navajo Nation Water Quality Project. Download the CSV file from
http://navajowater.org/export-raw-data/.
Water quality is a major issue on Native American reservations in the southwestern United States. The
prevalence of uranium mines and uranium mill accidents mean that much of the water in the Navajo Nation
is irradiated, and many homes are left without clean, drinkable water. Multiple environmental agencies
routinely sample water in the region and report on contaminants.


Problem 3 ->

The concentration of radiactive elements in a sample is measured in 
rate of atomic disintegrations per volume,
rather than mass per volume, as used for stable isotopes.
This is done by counting the number of atomic
disintegrations per minute and comparing it to the 
mass of the material involved. However, laboratory
environments and instruments used for detection 
create some number of atomic emissions on their own, so
background correction must be performed. 
Because this process involves sampling many times, and the
background can be inconsistent, 
resulting in over-correction, sometimes negative values are reported for the
concentration. For practical purposes
these values can be considered zero.
Mutate the dataset to replace the negative
values of Radium-228 with 0, then filter the dataset to remove
any sites with “Unknown Risk” for the EPA risk rating.
Visualize the distribution of Radium-228 
within each combination of EPA section and risk level. 
State your observations.


Answer:

```{r}
library(readr)
navajo<-read_csv("NavajoWaterExport.csv")
navajo<-mutate(navajo,`Amount of Radium228`=ifelse(`Amount of Radium228`<0.0,0,`Amount of Radium228`))
navajo$`Amount of Radium228`
```

```{r}
nv1<-filter(navajo,`US EPA Risk Rating`!="Unknown Risk")
"Unknown Risk" %in% nv1$`US EPA Risk Rating`
head(nv1)
```


```{r}
nv1_df <- summarise(group_by(nv1, `US EPA Risk Rating`, `Which EPA Section is This From?` ,`Amount of Radium228`))
ggplot(nv1_df) + 
  geom_histogram(aes(x=nv1_df$`Amount of Radium228`),bins=20) +
  facet_grid(nv1_df$`US EPA Risk Rating` ~nv1_df$`Which EPA Section is This From?`)+labs(x="Amount of Radium228")

```

```{r}
library(ggplot2)
#install.packages("maps")
#install.packages("mapproj")
```
Observations: In plot above for less risk in each section, 
amount of radium228 is less along with its frequency for each section.
For some risk there is high frequency of radium 228 in 
lower amount in every section. 
For more risk amount of radium is spread around 
larger spectrum with its frequency highest in section 2, 
that means more number of sites of section 2 come under more risk. 


Problem 4 ->
Create a map of the region 
(you may want to adjust the plotting limits to an appropriate “zoom” level)
showing the locations of the water sampling sites, 
along with the EPA risk and the concentration of Radium-228 for
each location (mapped to an appropriate aesthetic).

Answer:

```{r}
library(maps)
four_corners<-map_data("state",region = c("arizona","utah","new mexico","colorado"))
fc<-four_corners
#deg_dec_min, deg_min_sec 
navajo$Longitude<-measurements::conv_unit(navajo$Longitude,"deg_min_sec","dec_deg")
navajo$Latitude<-measurements::conv_unit(navajo$Latitude,"deg_min_sec","dec_deg")
nv2<-navajo
nv2$Longitude<-as.numeric(nv2$Longitude)
nv2$Latitude<-as.numeric(nv2$Latitude)
fc$long<-abs(fc$long)
ggplot(fc)+geom_polygon(mapping=aes(x=long,y=lat,group=group),fill="white",color="black")+geom_point(data=nv2,mapping=aes(x=Longitude,y=Latitude,size=(`Amount of Radium228`),color=`Which EPA Section is This From?`))+coord_quickmap()+xlim(100,NA)+ylim(31.25,NA)

```



Problems 5 ->
uses data from the US Department of Education’s Civil Rights Data Collection.
Download the zipped 2015-2016 data from
https://www2.ed.gov/about/offices/list/ocr/docs/crdc-2015-16.html. The
Public Use Data File User’s Manual should be included in the zipped files, 
or can be downloaded at the same location. 
Use it as a reference to help you understand the dataset.



We would like to investigate whether Black students
receive a disproportionate number of in-school suspensions.
Create a new data.frame or tibble with the following columns:
• The total number of students enrolled at each school
• The number of Black students enrolled at each school
• The total number of students who received one or 
more in-school suspension (including non-disabled
students and disabled students served by IDEA) at each school
• The number of Black students who received one or 
more in-school suspension (including non-disabled
students and disabled students served by IDEA) at each school
• The proportion of Black students at each 
school among all students at that school
• The proportion of Black students who 
received one or more in-school suspension among all suspended
students at that school
Plot the proportion of Black students at 
each school (on the x-axis) versus the proportion of suspended
students who are Black (on the y-axis). 
Include a smoothing line on the plot.
What do you observe in the plot?
Does the plot indicate an over- or under-representation of Black students
in in-school suspensions?
Calculate the overall proportion of Black students
across all schools and the overall proportion of suspended
students who are Black across all schools.
Are Black students over- or under-represented in in-school suspensions?

Answer:
```{r}
lea<-read_csv("D:/Spring 20 Sem 2/DMP/CRDC 2015-16 LEA Data.csv",na=c("-2","-5","-6","-7","-8","-9"))
school<-read_csv("D:/Spring 20 Sem 2/DMP/CRDC 2015-16 School Data.csv",na=c("-2","-5","-6","-7","-8","-9"))
head(lea)
dim(school)
head(school)
school1<-school
```


```{r}
school1$tot_stud=school1$TOT_ENR_M+school1$TOT_ENR_F
school1$tot_stud_black=school1$SCH_ENR_BL_M+school1$SCH_ENR_BL_F
school1$tot_suspension=school1$TOT_DISCWDIS_ISS_IDEA_M + school1$TOT_DISCWDIS_ISS_IDEA_F+school1$TOT_DISCWODIS_ISS_M + school1$TOT_DISCWODIS_ISS_F
school1$tot_suspension_black=school1$SCH_DISCWODIS_ISS_BL_M + school1$SCH_DISCWODIS_ISS_BL_F + school1$SCH_DISCWDIS_ISS_IDEA_BL_M +school1$SCH_DISCWDIS_ISS_IDEA_BL_F
school1$proportion_black=school1$tot_stud_black/school1$tot_stud
school1$susp_prop_black = school1$tot_suspension_black/school1$tot_suspension
x<-select(school1, tot_stud,tot_stud_black,tot_suspension,tot_suspension_black,proportion_black,susp_prop_black)
```


From the scatterplot below, we can see that proportion 
of suspended black students is more in schools 
where proportion of black students is less.
```{r}
x%>%sample_n(10000)%>%ggplot(aes(x=proportion_black,y=susp_prop_black))+geom_point()+geom_smooth()+xlab("Proportion of black students")+ylab("proportion of suspended black students")
```
In graph above we can observe under representation of black students.
Most of the suspended black students are within 
50% of proportion of black students.
Also as the proportion of black student increases 
proportion of suspended black students steadily decreases.
```{r}
tot_student<-sum(x$tot_stud,na.rm=TRUE)
tot_black<-sum(x$tot_stud_black,na.rm=TRUE)
tot_black_suspended<-sum(x$tot_suspension_black,na.rm=TRUE)
tot_stud_suspended<-sum(x$tot_suspension,na.rm=TRUE)
prop_black<-tot_black/tot_student
prop_black_suspended<-tot_black_suspended/tot_stud_suspended
cat("Overall black student proportion:",prop_black)
cat("Overall suspended black student proportion:",prop_black_suspended)
```

```
From above data, black students are under represented 
in school suspension as proportion of students suspended 
who are black holds smaller percentage of population.