---
title: "hw3-Rohit-Thakur"
author: "Rohit Thakur"
date: "2/9/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown
Problems 1–2 use data from the US Department of Education’s Civil Rights Data Collection. 
Download the zipped 2015-2016 data from
https://www2.ed.gov/about/offices/list/ocr/docs/crdc-2015-16.html. 
The Public Use Data File User’s Manual should be included in the zipped files,
or can be downloaded at the same location.
Use it as a reference to help you understand the dataset.
Read the documentation for the function read_csv() from the readr package, 
and use it to import the dataset into R. 
Check the User’s Manual for how missing values were reported, and handle them appropriately.
Treat all reserve codes as missing.

PART A

PROBLEM 1 ->

We would like to investigate whether Hispanic and Native American (American Indian / Alaska Native)
students are over- or under-represented in Gifted & Talented programs.
Create a new data.frame containing only schools with a Gifted & Talented program with the following
columns:
• The total number of students enrolled at each school
• The number of Hispanic students and Native American students at each school
• The total number of students in the school’s GT program
• The number of students in the GT program who are Hispanic or Native American
• The proportion of students at each school who are Hispanic or Native American among all students
• The proportion of students in the GT program who are Hispanic or Native American among students
in the GT program
Plot the proportion of Hispanic and Native American students at each school (on the x-axis) versus the
proportion of GT students who are Hispanic and Native American (on the y-axis). Include a smoothing line
on the plot.
What do you observe in the plot? 
Does the plot indicate an over- or under-representation of Hispanic and
Native American students in Gifted & Talented programs?
Calculate the overall proportion of Hispanic and 
Native American students across all schools and the overall
proportion of GT students who are Hispanic and Native American.
Are Hispanic and Native American students over- 
or under-represented in Gifted & Talented programs?

Answer:

```{r}
library(readr)
library(tidyr)
library(dplyr)
library(ggplot2)
school<-read.csv("D:/Spring 20 Sem 2/DMP/2015-16-crdc-data (1)/Data Files and Layouts/CRDC 2015-16 School Data.csv",na=c("-2","-5","-6","-7","-8","-9"))
#Filtering for schools containing GT program
school<-filter(school,school$SCH_GT_IND == "Yes") 
school1<-school
#Total Enrollment
school1$Tot_enr<-school1$TOT_ENR_M + school1$TOT_ENR_F
#Total students in gt program
school1$Tot_stud_gt<-school$TOT_GTENR_M + school$TOT_GTENR_F
#total hispanic and native americans in gt program
school1$Tot_HN_GT<-school1$SCH_GTENR_HI_M + 
  school1$SCH_GTENR_HI_F + 
  school1$SCH_GTENR_AM_M + school1$SCH_GTENR_AM_F
#total hispanic and native americans in school
school1$TOT_HN<-school1$SCH_ENR_HI_M +
  school1$SCH_ENR_HI_F + 
  school1$SCH_ENR_AM_F + school1$SCH_ENR_AM_M
#Proportion of hispanic or native americans amongst everyone in the school
school1$PROP_HN<-(school1$TOT_HN)/(school1$Tot_enr)
#Proportion of hispanic or native americans in gt program amongst evryone in the gt program
school1$PROP_HN_GT<-(school1$Tot_HN_GT)/(school1$Tot_stud_gt)
schgt<-select(school1,Tot_enr,Tot_stud_gt,Tot_HN_GT,TOT_HN,PROP_HN,PROP_HN_GT)
#New dataframe with required columns
schgt%>%sample_n(10000)%>%ggplot(aes(x=PROP_HN,y=PROP_HN_GT))+
  geom_point()+
  geom_smooth()+
  xlab("Proportion of Hispanic and Native") + 
  ylab("Proportion of Hispanic and Native in GT program")
```
In above graph by observing nature of the smooth line we can see that there is under representation of proportion of hispanic and native students in GT program.

```{r}
summarise(schgt,
pr_hn=sum(TOT_HN, na.rm=TRUE) / sum(Tot_enr, na.rm=TRUE),
pr_hn_gt=sum(Tot_HN_GT, na.rm=TRUE) / sum(Tot_stud_gt, na.rm=TRUE))
```
As we can see above overall proportion of hispanic and native american student in gifted
and talented program is less than that of overall proportion of hispanic and
native american students. Thus Hispanic and Native American students are under-represented
in Gifted & Talented programs.




PROBLEM 2->
We would like to investigate whether disabled 
students are disproportionately referred to law enforcement for discipline.
Create a new data.frame with the following columns:
• The total number of students enrolled at each school
• The number of disabled students (served by IDEA) at each school
• The total number of students who were referred to law enforcement
• The number of disabled students (served by IDEA) who were referred to law enforcement
• The proportion of disabled students (served by IDEA) at each school among of all students
• The proportion of students who were referred to law enforcement and are disabled (served by IDEA)
among all students referred to law enforcement
Plot the proportion of disabled students at each school 
(on the x-axis) versus the proportion of students were
referred to law enforcement and are disabled (on the y-axis). 
Include a smoothing line on the plot.
What do you observe in the plot? 
Does the plot indicate an over- or under-representation of disabled students
among students who are referred to law enforcement?
Calculate the overall proportion of disabled students 
across all schools and the overall proportion of students
who were referred to law enforcement and 
are disabled (served by IDEA) among all students referred to law
enforcement across all schools.
Are disabled students disproportionately referred to law enforcement?


Answer:


```{r}
school2<-read.csv("D:/Spring 20 Sem 2/DMP/2015-16-crdc-data (1)/Data Files and Layouts/CRDC 2015-16 School Data.csv",na=c("-2","-5","-6","-7","-8","-9"))
#CREATE A DATAFRAME WITH FOLLOWING

# The total number of students enrolled at each school
# The number of disabled students (served by IDEA) at each school
# The total number of students who were referred to law enforcement
# The number of disabled students (served by IDEA) who were referred to law enforcement
# The proportion of disabled students (served by IDEA) at each school among of all students
# The proportion of students who were referred to law enforcement and are disabled (served by IDEA)among all students referred to law enforcement

sch_disable <- transmute(school2,
enr_tot = TOT_ENR_M + TOT_ENR_F,
enr_dis_tot = TOT_IDEAENR_M + TOT_IDEAENR_F,
tot_law = TOT_DISCWODIS_REF_M + TOT_DISCWODIS_REF_F + TOT_DISCWDIS_REF_IDEA_F + TOT_DISCWDIS_REF_IDEA_M , 
tot_law_idea = TOT_DISCWDIS_REF_IDEA_F + TOT_DISCWDIS_REF_IDEA_M,
pr_disabled = enr_dis_tot/ enr_tot,
pr_law = tot_law_idea / tot_law)

#PLOT 
sch_disable%>%
sample_n(10000) %>%
ggplot(aes(x=pr_disabled, y=pr_law)) +
geom_point() +
geom_smooth(se=FALSE) +
geom_segment(aes(x=0, y=0, xend=1, yend=1), color="red") +
coord_fixed(x=c(0,1), y=c(0,1)) +
labs(x='Proportion of Disabled students at each school', y='Proportion of disabled students referred
     to law enforcements')
```
From the smooth line we can see over representation of proportion of disabled students who were referred to law enforcement. A lot of schools have below 25% percentage of disabled students. Similarly students referred to law enforcements hav maximum population density below 50%.


From the proportions below disabled students are tend to be over represented among disabled students
referred to law enforcement.
```{r}
#Calculating overall proportions
summarise(sch_disable,
ovr_pr_dis=sum(enr_dis_tot, na.rm=TRUE) / sum(enr_tot, na.rm=TRUE),
ovr_pr_law=sum(tot_law_idea, na.rm=TRUE) / sum(tot_law, na.rm=TRUE))

```



PART B
Problems 3–5 uses a subset of the DBLP database of bibliographic 
information on major computer science
journals and proceedings, available from 
https://data.mendeley.com/datasets/3p9w84t5mr. 
The dataset has been processed to include predictions 
of the author’s genders using the open-source Genderize API.
The processed data has been made available in the form of SQL scripts 
that import the data into a MySQL database.
We are primarily interested in the “general” and “authors” tables created by the “main.sql” and
“authors.sql” scripts, respectively.
```{r}
library(dbplyr)
library(DBI)
library(RMySQL)
con<-dbConnect(MySQL(),user="root",password="password",dbname="dblp",host="localhost")
dblp_main<-tbl(con,"general")
dbListTables(con)
dblp_authors<-tbl(con,"authors")
dblp_authors<-as.data.frame(dblp_authors)
dblp_main<-as.data.frame(dblp_main)
```

PROBLEM 3
Filter the data to include only the authors for whom a gender was 
predicted as ‘male’ or ‘female’ with a
probability of 0.90 or greater, and 
then create a bar plot showing the total number of distinct male 
and female authors published each year. 
Comment on the visualization


Answer:

From the graph below there is huge difference between male and female authors 
getting published each year. Male authors are substantially getting increased publications
each year. Female authors published are less than 5000 over all the years.
```{r}
dblp_authors_filt<-dblp_authors%>%filter((gender=="M"|gender=="F") & prob>=0.90)
dblp_authors_filt %>%
  inner_join(dblp_main,by=c("k"="k"))%>%
  distinct()%>%
  ggplot()+geom_bar(aes(x=factor(year),fill=gender),position="dodge")+
  xlab("Distinct male and female authors published each year")+
  ylab("total number") + theme(axis.text.x = element_text(angle = 90))
```


PROBLEM 4
Still including only the authors for whom a gender was
predicted with a probability of 0.90 or greater,
create a stacked bar plot showing the 
proportions of distinct male authors vs. distinct female authors published each
year. 
(The stacked bars for each year will sum to one.) 
Comment on the visualization.


Answer:

From the graph below we can see that proportions of female authors are steadily increasing.
For the first few years there were no female candidates. Male candidates are in huge amount compared to female authors.
```{r}
dblp_authors_filt%>%
  inner_join(dblp_main,by=c("k"="k"))%>%
  distinct()%>%
  ggplot()+geom_bar(aes(x=factor(year),fill=gender),position="fill")+
  xlab("Distinct male and female authors published each year")+
  ylab("proportion") + theme(axis.text.x = element_text(angle = 90))
```





PROBLEM 5

Still including only the authors for whom a gender was predicted 
with a probability of 0.90 or greater, 
create a faceted bar plots showing the 
proportions of female first authorships for each year for each domain (CS, DE,
SE, and TH). 
(If a conference belongs to multiple domains, you may include it for both of them.) 
Comment on the visualization. 
Which domains have the highest and lowest representation of papers with female first
authors?



Answer:
In the graph below CS has highest representation of female first authors and 
SE has lowest representation of female first authors. 
Also we can see steady increase in female first author in cs unlike other types of conferences.
SE also has more number of female authors than before.
Some years are missing as there are no first female authors in those years for that type. 
```{r}
trial_1 <-
dblp_authors_filt %>%
  inner_join(dblp_main,by=c("k"="k"))%>%
  distinct()%>%filter(pos==0)
trial_2 <- trial_1 %>% gather(cs,de,se,th,key="type",value="count")%>%
  filter(count==1)
trial_3 <- trial_2 %>% group_by(year, type, gender) %>% tally()
trial_4 <- summarise(group_by(trial_3, year, type, gender, n))
trial_5_sum_n <- summarise(group_by(trial_4, year), sum(n))
trial_6 <- merge(trial_4, trial_5_sum_n, by = "year")
trial_6 <- trial_6 %>% filter(gender == "F")
trial_7 <- summarise(group_by(trial_6, year, type, gender, n), proportion = n / `sum(n)`)
ggplot(trial_7)+geom_bar(aes(x=factor(year), y=proportion), stat = "identity")+
  facet_wrap(~type) +
  xlab("Years")+ 
  ylab("Proportion of female with first authorship")+
  theme(axis.text.x = element_text(angle = 90))
```