---
title: "HW4-Rohit-Thakur"
author: "Rohit Thakur"
date: "2/24/2020"
output: pdf_document
---
```{r}
load("D:/Spring 20 Sem 2/DMP/ICPSR_31721/DS0001/31721-0001-Data.rda")
df<-da31721.0001
library(tidyr)
library(dplyr)
library(ggplot2)
```
PROBLEM 1
```{r}
df1 <- 
  mutate(df,Q6=ifelse(
      (df$Q6=="(1) Man" & df$Q5=="(2) Female"),"transmen",
      ifelse((df$Q6=="(2) Woman" & df$Q5=="(1) Male"),"transwomen",
      ifelse((df$Q6=="(4) Androgynous" | df$Q6=="(6) Gender Queer"),"non-binary", 
      df$Q6))))

df1<- filter(df1,(Q6=="transmen" | Q6=="transwomen" | Q6=="non-binary"))
df1%>%ggplot()+
  geom_bar(aes(x=as.factor(Q6)))+
  xlab("Category")+
  ylab("count")+
  ggtitle("Count for each gender category")
df1<- filter(df1,Q88!="(10) Missing") 
df1%>%dplyr::group_by(Q6)%>%
  summarise(y = sum(Q88=="(1) Yes"), n= sum(Q88 == "(2) No"))%>%
  mutate(prop=y/(y+n))%>%
ggplot() + geom_col(aes(x=Q6,y=prop))+
  xlab("Gender Category")+ylab("Proportion") + 
  ggtitle("Proportion of participants who have ever been homeless")
#Proportion of homeless participants
df1%>%summarize(proportion_homeless=mean(Q88 == "(1) Yes"))
```
participants in our survey have experienced homelessness more than general US population.
Problem 2
```{r}
df2<-df
df2<-rename(df2,African_American=D9_1,Caucasian=D9_2,Hispanic_Latino=D9_3, Native_American=D9_4)
df2<-df2%>%
  gather(`African_American`,
         `Caucasian`,
         `Hispanic_Latino`,
         `Native_American`,
          key="Race",
          value="value")%>%
  filter(value=="(1) Selected")
summary(df2$Race)
df2<-df2%>% filter(Q88!="(10) Missing")
summary(df2$Race)
df2%>%
  group_by(Race)%>%
  summarise(y = sum(Q88=="(1) Yes"), n= sum(Q88 == "(2) No"))%>%
  mutate(prop=y/(y+n))%>%
  ggplot() + geom_col(aes(x=Race,y=prop))+
  xlab("Gender Category")+ylab("Proportion") +
  ggtitle("Proportion of participants who have ever been homeless")
df2%>%summarize(proportion_homeless=mean(Q88 == "(1) Yes"))
```
From above proportion 25.96% participants from our study experience homelessness
in their lifetime. This is greater than national US average.

```{r}
df3<-df
df3<-df3%>%mutate(Q133=ifelse(is.na(df3$Q133),"Missing",df3$Q133))
summary(df3$Q133)
#Mapped NAs to missing and yes response to 1 and no response to 2
df3%>%summarize(proportion_suicide_attempt=mean(df3$Q133=="1"))
summary(df3$Q119)
df4<-df3
df4<-df4[!is.na(df4$Q119),]
summary(df4$Q119)
df4<-filter(df4,Q119!="(5) Not applicable to me")
df4<-df4[!is.na(df4$Q131),]
df4%>%group_by(Q119)%>%
  summarise(y = sum(Q131=="(1) Yes"), n= sum(Q131 == "(2) No"))%>%
  mutate(prop=y/(y+n))%>%
  ggplot()+geom_col(aes(x=Q119,y=prop))+
  xlab("Suicidal thoughts")+ylab("Proportion")+
  ggtitle("proportions of participants thought about killing themselves vs familial support")   
```
From above proportion of participants who have attempted suicide in Virginia in this survey
is 25.14% which is below national average for trans people but much higher than average
proportion of general population. 


From above graph for participants who have thought of attempting suicide vs familial support levels
we can conclude that there is higher percentage of participants who have positive suicidal thoughts 
overall in all familial support categories. But family support surely reduces percentage of population
of participants that are having positive suicidal thoughts. Highest percentage of participants that have
positive suicidal thoughts can be seen in participants with no supoortive family. This proportion get
reduced as the family support level increases.



Problem 4
```{r}
library(mlbench)
library(modelr)
```
```{r}
data("PimaIndiansDiabetes2")
#First we will select pregnant column for mass. 
PimaIndiansDiabetes2%>%ggplot(aes(x=pregnant,y=mass))+geom_point()+geom_smooth()

#Glucose vs mass
PimaIndiansDiabetes2%>%ggplot(aes(x=glucose,y=mass))+geom_point()+geom_smooth()

#Pressure vs mass
PimaIndiansDiabetes2%>%ggplot(aes(x=log(pressure),y=mass))+geom_point()+geom_smooth()
#from above graph there is a linear relationship with pressure and mass. Log og predictor variable
#better helps in visualizing this relationship.
#Triceps vs mass

PimaIndiansDiabetes2%>%ggplot(aes(x=triceps,y=log2(mass)))+geom_point()+geom_smooth()
#From above we can see a linear relationship between log of response variable and triceps
#log of response variable better depicts above relationship.

#Insulin vs mass
PimaIndiansDiabetes2%>%ggplot(aes(x=insulin,y=mass))+geom_point()+geom_smooth()


#pedigree vs mass
PimaIndiansDiabetes2%>%ggplot(aes(x=pedigree,y=mass))+geom_point()+geom_smooth()

#age vs mass
PimaIndiansDiabetes2%>%ggplot(aes(x=age,y=mass))+geom_point()+geom_smooth()

#diabetes vs mass
PimaIndiansDiabetes2%>%ggplot(aes(x=diabetes,y=mass))+geom_boxplot()


```
From above plots we can see that there is a linear relationship between (pressure,mass) and 
(triceps,mass). Thus we select these two columns as predictor variables. There are simply not 
enough categories to decide whether relationship between diabetes and mass is linear.
Now fitting linear model for above two variables:

```{r}
fit_1<-lm(mass~pressure+triceps,data=PimaIndiansDiabetes2)
summary(fit_1)
```
Problem 5
```{r}
#residuals for predictor variable pressure
PimaIndiansDiabetes2%>%
add_residuals(fit_1, "residuals") %>%
ggplot(aes(x=pressure)) + geom_point(aes(y=residuals))


#residuals for predictor variable triceps

PimaIndiansDiabetes2%>%
add_residuals(fit_1, "residuals") %>%
ggplot(aes(x=triceps)) + geom_point(aes(y=residuals))

#From residual plots we can observe they are randomly dispersed around horizontal axis.
#This proves linear regression model is appropriate for this data.

#relationship between the residuals and the other potential predictor variables that are not
#currently in model

#for diabetes
PimaIndiansDiabetes2 %>%
add_residuals(fit_1, "residuals") %>%
ggplot(aes(diabetes, residuals)) + geom_boxplot()

#for pregnant

PimaIndiansDiabetes2 %>%
add_residuals(fit_1, "residuals") %>%
ggplot(aes(pregnant, residuals)) + geom_point() + geom_smooth()

#for glucose

PimaIndiansDiabetes2 %>%
add_residuals(fit_1, "residuals") %>%
ggplot(aes(glucose, residuals)) + geom_point()

#for insulin

PimaIndiansDiabetes2 %>%
add_residuals(fit_1, "residuals") %>%
ggplot(aes(insulin, log2(residuals))) + geom_point() + geom_smooth()

#for pedigree

PimaIndiansDiabetes2 %>%
add_residuals(fit_1, "residuals") %>%
ggplot(aes(pedigree, residuals)) + geom_point() + geom_smooth()

#for age

PimaIndiansDiabetes2 %>%
add_residuals(fit_1, "residuals") %>%
ggplot(aes(age, residuals)) + geom_point() + geom_smooth()

fit_3<-lm(mass~pressure+triceps+pregnant,data=PimaIndiansDiabetes2)
summary(fit_3)

# we will find rmse for two ,three, four variables

fit_4<-lm(mass~pressure+triceps+diabetes+pregnant,data=PimaIndiansDiabetes2)
summary(fit_4)
rmse(fit_1,PimaIndiansDiabetes2)
rmse(fit_3,PimaIndiansDiabetes2)
rmse(fit_4,PimaIndiansDiabetes2)
# we can see rmse is decreasing as we add diabetes and pregnant predictor variable compared to 
#two variable model.

#visualizing rmse
fits_rmse <- tibble(nvar = 1:3,
rmse = c(rmse(fit_1,PimaIndiansDiabetes2),
rmse(fit_3,PimaIndiansDiabetes2),
rmse(fit_4,PimaIndiansDiabetes2)))
ggplot(fits_rmse) + geom_line(aes(x=nvar, y=rmse)) +xlab("Variables")
```
From above residual plots we cannot see any definitive systematic pattern for variables. But
for diabetes and pregnant variables there is some pattern.
we'll add this predictor variables to our model.
After calculating rmse for three and four variables model we compare it
and find value to be lower.

Plotting RMSE we get linear decrease in rmse for these three fits.