---
title: "HW5-Rohit Thakur"
author: "Rohit Thakur"
date: "3/19/2020"
output: pdf_document
---
Problem 1:
Author: Maanasa Kaza
Source: https://www.kaggle.com/carrie1/ecommerce-data/data#
```{r}
library(tidyr)
library(readr)
library(dplyr)
library(ggplot2)
library(forcats)
library(lubridate)
df<-read.csv("D:/Spring 20 Sem 2/DMP/data.csv")
head(df,n=10)
```
Problem 2
```{r}
#Putting into tidy format
df<-separate(df,col="InvoiceDate",into=c("Date","Time"),sep=" ")
head(df)
df<-filter(df,Description!="")
df<-na.omit(df)
df<-filter(df,UnitPrice > 0)
df<-filter(df,Quantity>0)
head(df, n=10)
df<-mutate(df,Time=hm(Time))
df1<-df
df1<-df1%>%mutate(Expenditure=Quantity*UnitPrice)
```
```{r}
df2 <- summarise(group_by(df1, Description), Sales=sum(Expenditure))
df2<-df2%>% arrange(desc(Sales))
df3<-df2[1:5,]
ggplot()+geom_col(aes(x=reorder(df3$Description, df3$Sales),y=df3$Sales)
                  ,fill='chocolate',width=0.7)+
  coord_flip()+xlab("Product Description")+ylab("Sales")
```
```{r}
df4<-df1%>%distinct(InvoiceNo,.keep_all=TRUE)
df4<-separate(df4,col="Date",into=c("mm","dd","yyyy"),sep="/")
df4<-df4%>%filter(yyyy=="2011")
df4<-df4%>%group_by(mm)%>%summarise(Orders=n())

df4$mm <- df4$mm%>%factor(levels = c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12))
df4<- df4%>%
  mutate(mm=recode(mm, "1"="Jan","2"="Feb","3"="Mar","4"="Apr","5"="May","6"="June",
                       "7"="July","8"="Aug","9"="Sept","10"="Oct","11"="Nov","12"="Dec"))
df4%>%ggplot()+geom_line(aes(x=mm,y=Orders,group=1))+xlab("Month")
```

```{r}
  
df5<-select(df1,Date,Time,Expenditure)
  df5$day <- wday(as.Date(df5$Date,'%m/%d/%Y'), label = TRUE)

  
  df5<-group_by(df5,hour(Time),day)%>%summarise(total=sum(Expenditure))
  plot<- ggplot(data = df5 ,
                        mapping = aes(x = `hour(Time)`, y = total, group=day)) +
  geom_line(stat = 'identity', aes(color = day)) +
  labs(x = 'Hour', y = 'Sales')
  print(plot)
```


Problem 3
```{r}
library(mlbench)
library(modelr)
library(purrr)
data(PimaIndiansDiabetes2)
cv<-function(formula,data,folds){
  set.seed(1)
  data_cv <- crossv_kfold(data,folds)
  data_cv <- data_cv%>%
              mutate(fit = map(train,
                 ~ lm(formula,data=data)))
  data_cv <- data_cv %>%
            mutate(rmse_test = map2_dbl(fit, test, ~ rmse(.x,.y)))
  mean_error<-mean(data_cv$rmse_test)
  return(mean_error)
}

cv(mass~pressure+triceps+diabetes+pregnant,PimaIndiansDiabetes2,5)
```




Problem 4

First fit for each predictor variable
```{r}
rmse_triceps<-cv(mass~triceps,PimaIndiansDiabetes2,5)
print(rmse_triceps)
rmse_diabetes<-cv(mass~diabetes,PimaIndiansDiabetes2,5)
print(rmse_diabetes)
rmse_pressure<-cv(mass~pressure,PimaIndiansDiabetes2,5)
print(rmse_pressure)
rmse_insulin<-cv(mass~insulin,PimaIndiansDiabetes2,5)
print(rmse_insulin)
rmse_glucose<-cv(mass~glucose,PimaIndiansDiabetes2,5)
print(rmse_glucose)
```
triceps is the best predictor with rmse 5.222
```{r}
rmse_diabetes2<-cv(mass~triceps+diabetes,PimaIndiansDiabetes2,5)
print(rmse_diabetes2)
rmse_pressure2<-cv(mass~triceps+pressure,PimaIndiansDiabetes2,5)
print(rmse_pressure2)
rmse_insulin2<-cv(mass~triceps+insulin,PimaIndiansDiabetes2,5)
print(rmse_insulin2)
rmse_glucose2<-cv(mass~triceps+glucose,PimaIndiansDiabetes2,5)
print(rmse_glucose2)
```
Pressure is best predictor with rmse 5.09 we will add it in our model

```{r}
rmse_diabetes3<-cv(mass~triceps+pressure+diabetes,PimaIndiansDiabetes2,5)
print(rmse_diabetes3)
rmse_insulin3<-cv(mass~triceps+pressure+insulin,PimaIndiansDiabetes2,5)
print(rmse_insulin3)
rmse_glucose3<-cv(mass~triceps+pressure+glucose,PimaIndiansDiabetes2,5)
print(rmse_glucose3)
```
we will further add diabetes in our model because it reduces rmse to 5.02.
```{r}
rmse_insulin4<-cv(mass~triceps+pressure+diabetes+insulin,PimaIndiansDiabetes2,5)
print(rmse_insulin4)
rmse_glucose4<-cv(mass~triceps+pressure+diabetes+glucose,PimaIndiansDiabetes2,5)
print(rmse_glucose4)
```
Addition of predictor variables further does not improve
model. Thus final model will be mass~triceps+pressure+diabetes
with rmse 5.02
```{r}
fits_rmse <- tibble(nvar = 1:4,
                    rmse = c(rmse_triceps,rmse_pressure2,rmse_diabetes3,rmse_glucose4))
ggplot(fits_rmse) + geom_line(aes(x=nvar, y=rmse))
```

In homework 4 proposed model for response variable mass
was using 4 predictor variables(pressure,triceps,diabetes,pregnant)
it had rmse 4.97 while current model 
with predictor variables(triceps,pressure,diabetes) gives rmse 5.02 which is
slightly worse than previous model.




PROBLEM 5

Final cross validated rmse for best model we found in problem 4
CANNOT be reported as good measure of rmse on new data. 
One of the reason is that best model for this data may not provide good 
estimate for new set of data.
For RMSE to be a good estimate , we should calculate it on set of test data
so as to make sure that it was not used to select the best model.